Seismic parameter estimation and the Canadian crust

### 2013-09-08 paper/draftBP.tex
thwat
disucssion

\begin{table}
  \begin{tabular}{ l l l l l l }
    \cline{3-6}
    & & \multicolumn{2}{ c }{Correlation} & \multicolumn{2}{ c }{Mean Difference} \\
    \hline
    Study Authors &   Num. of Stns &  $H$ & $R$  &$H$(km)& $R$ \\
    \hline
    Thompson et. al. (2010)   & 29 & 0.97 & 0.70 & 0.49 & 0.012 \\
    Darbyshire et. al. (2007) & 10 & 0.95 & 0.43 & 1.21 & 0.033 \\
    Eaton et. al. (2006)      & 26 & 0.90 & 0.61 & 0.79 & 0.025 \\
    \hline
  \end{tabular}
  \caption{Comparison of $R$ and $H$ estimates with three published studies}
\label{table:comparison}

@@

%% -----------------------------------------------%%
\section{Discussion}
\subsection{Canada}

\begin{itemize}

\item Quiet zone in discontinuity data

\item Compare data to Crust 2.0. What do the differences mean.
  \begin{itemize}
  \item Lith 5.0 data?
  \end{itemize}

\item Discuss spatial trends -> high R -> Low R moving North West through Ontario
  \begin{itemize}
  \item this is part of larger trend of changing R values as we move from Churchill to Grenville
  \item in the literature? Yes. See notes on lithoprobe studies
  \end{itemize}

\item Secular variation in NE Churchill from thinner crusts to thicker as we move East
  \begin{itemize}
  \item in the literature? See Lith 5.0: A revised crustal model for Canada based on lithoprobe results.
  \end{itemize}

 \item Explanation of results

 \item Canada H and Vp/Vs

 \item Lith5.0 (Perry, 2002) estimates for Canadian crustal averages is 38 km. The Crust
   2.0 average is 38 km as well, however this is expected since Crust2.0 data for
   Canada is based on the same data as that of Lith5.0 .... is it definately?

 \item Make some comparison to global average?

@@

@@

Perry, H. K. C., D. W. S. Eaton, A. M. Forte (2002) LITH5.0: a revised crustal model for Canada based on Lithoprobe results,  J. Geophys. Int., 150, 285-294.

>>>>

\begin{table}
  \begin{tabular}{ l l l l l l l}
    \cline{4-7}
    & & \multicolumn{2}{ c }{Correlation Coeff.} & \multicolumn{2}{ c }{RMS difference} \\
    \hline
    Study Authors             & Region         & Num. of Stns &  $H$ & $R$  &$H$(km)& $R$ \\
    \hline
    Thompson et. al. (2010)   & Northern Canadian Sheild & 29 & 0.97 & 0.70 & 0.78 & 0.017 \\
    Darbyshire et. al. (2007) & Superior Province        & 10 & 0.95 & 0.43 & 1.71 & 0.037 \\
    Eaton et. al. (2006)      & Grenville Orogen         & 26 & 0.90 & 0.61 & 1.03 & 0.033 \\
    \hline
  \end{tabular}
  \caption{Comparison of $R$ and $H$ estimates with three published studies}
\label{table:comparison}

@@

%% -----------------------------------------------%%
\section{Discussion}

Mean crustal composition is an important parameter in the literature studying the continental crust. Mean composition should not be confused as an estimate of real physical crustal lithological and mineral assemblages. Rather, mean crustal composition is a representative quantity that is useful for constraining crustal genesis and evolution in space and time. Early attempts at estimating mean crustal composition were largely accomplished through extrapolations of upper crustal rocks [Rudnick and Gao, 2003]. After the plate-tectonic revolution, Taylor [1967, 1977], developed the 'island arc' or 'andesite' model which posited that bulk crustal composition should be equivalent to the average convergent-margin andesite. Several problems with this theory, such as the fact that island arcs are more basaltic than andesitic and therefore produce basaltic crustal accretions and cannot account for the intermediate composition of post-Archean crust, led to the introduction of seismic methods to estimate mid and lower crustal composition [Christensen and Mooney, 1995; Rudnick and Fountain, 1995]. This study differs from these later efforts in that it follows more closely to the complete crustal seismic analysis in the work of Smithson et. al. [1981] and Zandt and Ammon [1995].

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{lithology}
  \caption{$V_P$ and $V_S$ estimates for common crust lithologies and for major Canadian-Shield provinces}
  \label{fig:lith}
\end{figure}

The mean seismic velocity ratio computed in this study is combined with values for $V_P$ from the Crust 1.0 model to enable estimates for $V_P$ and $V_S$ to be used to estimate mean crustal composition. Figure \ref{fig:lith} shows mineral assemblages and weighted Canadian Shield estimates plotted against $V_P$ and $V_S$. Mineral assemblages are chosen as useful representations of common crustal rocks and, specifically, for crust in Shield-like regions. Examples of common shield lithologies can be seen in the 30km of paleo-depth represented in the Kapuskasing uplift in the Superior province [Percival and West, 1994]. This uplift is largely comprised of Archean granite-greenstone terranes dominated by granitoid and gneissic rocks, and contains supracrustal assemblages generally metamorphosed to greenschist to amphibolite facies (hence the inclusion of hornblendite and amphibolite) and early plutonic structures comprised largely of tonalites. Tonalites and granodiorites are also part of the TGG suite (for tonalite-trondhjemite-granodiorite) which dominate Archaean cratons [Rudnick, 1995]. Seismic velocity values for these lithologies are taken from the work of Christensen [1996] and are selected at average crustal pressures of 600 MPa. Lines of constant Poisson's Ratio are plotted for reference.

The Slave, Superior and Churchill provinces share similar mean crustal compositions, with the Slave province being slightly more tonalitic in composition. These three provinces cluster around the average for the Canadian Shield. The Grenville Province on the other hand differs significantly in mean composition. One interpretation is that the crust in the Grenville province has a mean crustal composition corrisponding to a more mafic lower crustal composition than other Shield provinces. This interpretation consistent with a current model for the tectonic evolution of the central region of the Grenville Province that

@@

Christensen, N. I., W. D. Mooney (1995) Seismic velocity structure and composition of the continental crust: A global view, J. Geophys. Res., 100, 9761-9788.

@@

Percival, J. A., Gordon F. W. (1994), The Kapuskasing uplift: a geological and geophysical synthesis, Canadian. J. of Earth Sciences, 31, 1256-1286.

Perry, H. K. C., D. W. S. Eaton, A. M. Forte (2002), LITH5.0: a revised crustal model for Canada based on Lithoprobe results,  J. Geophys. Int., 150, 285-294.

Rudnick, R. L., D. M. Fountain (1995), Nature and composition of the continental crust: A lower crustal perspective, Rev. Geophys., 33(3), 267–309, doi:10.1029/95RG01302.

Rudnick, R. L., S. Gao (2003), Composition of the Continental Crust, Treatise on Geochem., 3, 1-64.

Smithson, S. B., R. A. Johnson, Y. K. Wong (1981), Mean crustal velocity: a critical parameter for interpreting crustal structure and crustal growth, Earth and Planetary Science Letters, 53, 323-332.

Taylor, S. R. (1967), The origin and growth of continents, Tectonophysics, 4, 17–34.

Taylor, S. R. (1977), Island arc models and the composition of the continental crust, in Island Arcs, Deep Sea Trenches and Back-Arc Basins, Maurice Ewing Ser., vol. 1, edited by M. Talwani and W. C. Pitman III, pp. 325–335, AGU, Washington, D. C., doi:10.1029/ME001p0325.
